---
title: "Design and run Workflows" 
author: "Author: Daniela Cassol (danielac@ucr.edu) and Thomas Girke (thomas.girke@ucr.edu)"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`" 
output:
  BiocStyle::html_document:
    toc_float: true
    code_folding: show
  BiocStyle::pdf_document: default
package: systemPipeR
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{systemPipeR: Workflow design and reporting generation environment}
  %\VignetteEngine{knitr::rmarkdown}
fontsize: 14pt
bibliography: bibtex.bib
editor_options: 
  chunk_output_type: console
type: docs
weight: 3
---

<!--
- Compile from command-line
Rscript -e "rmarkdown::render('systemPipeR.Rmd', c('BiocStyle::html_document'), clean=F); knitr::knit('systemPipeR.Rmd', tangle=TRUE)"; Rscript ../md2jekyll.R systemPipeR.knit.md 2; Rscript -e "rmarkdown::render('systemPipeR.Rmd', c('BiocStyle::pdf_document'))"
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
options(width=80, max.print=1000)
knitr::opts_chunk$set(
    eval=as.logical(Sys.getenv("KNITR_EVAL", "TRUE")),
    cache=as.logical(Sys.getenv("KNITR_CACHE", "TRUE")), 
    tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
    library(systemPipeR)
})
```

```{r cleaning2, eval=TRUE, include=FALSE}
if (file.exists(".SPRproject")) unlink(".SPRproject", recursive = TRUE)
## NOTE: Removing previous project create in the quick starts section
```

# About this section

In this section, we will discuss following topics: 

- How to create SPR data analysis projects.
- How to build workflow step by step interactively or use use a template as starting point.
- After step design, how to run a workflow.
- After workflow finished running, how we can check the status, visualize the workflow.
- Different options for managing the workflow, _e.g._ resume, restart, overwrite a 
  SPR project. 
- How to explore the workflow object (methods).
- Finally, how to generate some data analysis reports.


# Project initialization

To create a workflow within _`systemPipeR`_, we can start by defining an empty
container and checking the directory structure:

```{r SPRproject1, eval=TRUE}
sal <- SPRproject(projPath = getwd(), overwrite = TRUE) 
```

Internally, `SPRproject` function will create a hidden folder called `.SPRproject`,
by default, to store all the log files.
A `YAML` file, here called `SYSargsList.yml`, has been created, which initially
contains the basic location of the project structure; however, every time the 
workflow object `sal` is updated in R, the new information will also be store in this 
flat-file database for easy recovery.
If you desire different names for the logs folder and the `YAML` file, these can 
be modified as follows:

```{r SPRproject_logs, eval=FALSE}
sal <- SPRproject(logs.dir= ".SPRproject", sys.file=".SPRproject/SYSargsList.yml") 
```

Also, this function will check and/or create the basic folder structure if missing, 
which means `data`, `param`, and `results` folder, as described [here](https://systempipe.org/sp/spr/gettingstarted/#directory-structure). 
If the user wants to use a different names for these directories, can be specified 
as follows:

```{r SPRproject_dir, eval=FALSE}
sal <- SPRproject(data = "data", param = "param", results = "results") 
```

It is possible to separate all the R objects created within the workflow analysis 
from the current environment. `SPRproject` function provides the option to create 
a new environment, and in this way, it is not overwriting any object you may want
to have at your current section. 

```{r SPRproject_env, eval=FALSE}
sal <- SPRproject(envir = new.env()) 
```

In this stage, the object `sal` is a empty container, except for the project information. The project information can be accessed by the `projectInfo` method:

```{r projectInfo, eval=TRUE}
sal
projectInfo(sal)
```

Also, the `length` function will return how many steps this workflow contains, 
and in this case, it is empty, as follow:

```{r length, eval=TRUE}
length(sal)
```

# Workflow Design 

_`systemPipeR`_ workflows can be designed and built from start to finish with a single command, importing from an R Markdown file or stepwise in interactive mode from the R console. 
In the [next section](#appendstep), we will demonstrate how to build the workflow in an interactive mode, and in the [following section](#importWF), we will show how to build from a file. 

New workflows are constructed, or existing ones modified, by connecting each step
via `appendStep` method. Each `SYSargsList` instance contains instructions needed 
for processing a set of input files with a specific command-line and the paths to 
the corresponding outfiles generated.

The constructor function `Linewise` is used to build the R code-based step. 
For more details about this S4 class container, see [here](#linewise). 

## Build workflow interactive {#appendstep}

This tutorial shows a straightforward example for describing and explaining all main features available within systemPipeR to design, build, manage, run, and visualize the workflow. In summary, we are exporting a dataset to multiple files, compressing and decompressing each one of the files, importing to R, and finally performing a statistical analysis. 

In the previous section, we initialize the project by building the `sal` object.
Until this moment, the container has no steps:

```{r sal_check, eval=TRUE}
sal
```

In the next subsection, we will discuss how to populate the object created with the first step in the
workflow interactively.



# Running the workflow

For running the workflow, `runWF` function will execute all the command lines 
store in the workflow container.

```{r runWF, eval=FALSE}
sal <- runWF(sal)
```

This essential function allows the user to choose one or multiple steps to be 
executed using the `steps` argument. However, it is necessary to follow the 
workflow dependency graph. If a selected step depends on a previous step(s) that
was not executed, the execution will fail. 

```{r runWF_error, eval=FALSE}
sal <- runWF(sal, steps = c(1,3))
```

Also, it allows forcing the execution of the steps, even if the status of the 
step is `'Success'` and all the expected `outfiles` exists. 
Another feature of the `runWF` function is ignoring all the warnings 
and errors and running the workflow by the arguments `warning.stop` and 
`error.stop`, respectively.

```{r runWF_force, eval=FALSE}
sal <- runWF(sal, force = TRUE, warning.stop = FALSE, error.stop = TRUE)
```

When the project was initialized by `SPRproject` function, it was created an 
environment for all objects created during the workflow execution. This 
environment can be accessed as follows:

```{r runWF_env, eval=FALSE}
viewEnvir(sal)
```

The workflow execution allows to save this environment for future recovery:

```{r runWF_saveenv, eval=FALSE}
sal <- runWF(sal, saveEnv = TRUE)
```

## Workflow status

To check the summary of the workflow, we can use:

```{r show_statusWF, eval=TRUE}
sal
```

To access more details about the workflow instances, we can use the `statusWF` method:

```{r statusWF, eval=TRUE}
statusWF(sal)
```

## Parallelization on clusters

This section of the tutorial provides an introduction to the usage of the 
_`systemPipeR`_ features on a cluster.

Alternatively, the computation can be greatly accelerated by processing many files 
in parallel using several compute nodes of a cluster, where a scheduling/queuing
system is used for load balancing. 

The `resources` list object provides the number of independent parallel cluster 
processes defined under the `Njobs` element in the list. The following example 
will run 18 processes in parallel using each 4 CPU cores. 
If the resources available on a cluster allow running all 18 processes at the 
same time, then the shown sample submission will utilize in a total of 72 CPU cores.

Note, `runWF` can be used with most queueing systems as it is based on utilities 
from the `batchtools` package, which supports the use of template files (_`*.tmpl`_)
for defining the run parameters of different schedulers. To run the following 
code, one needs to have both a `conffile` (see _`.batchtools.conf.R`_ samples [here](https://mllg.github.io/batchtools/)) 
and a `template` file (see _`*.tmpl`_ samples [here](https://github.com/mllg/batchtools/tree/master/inst/templates)) 
for the queueing available on a system. The following example uses the sample 
`conffile` and `template` files for the Slurm scheduler provided by this package. 

The resources can be appended when the step is generated, or it is possible to 
add these resources later, as the following example using the `addResources` 
function:

```{r runWF_cluster, eval=FALSE}
resources <- list(conffile=".batchtools.conf.R",
                  template="batchtools.slurm.tmpl", 
                  Njobs=18, 
                  walltime=120,##minutes
                  ntasks=1,
                  ncpus=4, 
                  memory=1024,##Mb
                  partition = "short"
                  )
sal <- addResources(sal, c("hisat2_mapping"), resources = resources)
sal <- runWF(sal)
```

Note: The example is submitting the jog to `short` partition. If you desire to 
use a different partition, please adjust accordingly.

# Visualize workflow

_`systemPipeR`_ workflows instances can be visualized with the `plotWF` function.

This function will make a plot of selected workflow instance and the following 
information is displayed on the plot:

    - Workflow structure (dependency graphs between different steps); 
    - Workflow step status, *e.g.* `Success`, `Error`, `Pending`, `Warnings`; 
    - Sample status and statistics; 
    - Workflow timing: running duration time. 

If no argument is provided, the basic plot will automatically detect width, 
height, layout, plot method, branches, _etc_. 

```{r, eval=TRUE}
plotWF(sal, show_legend = TRUE, width = "80%", rstudio = TRUE)
```

For more details about the `plotWF` function, please see [here](#plotWF).

# Technical report

_`systemPipeR`_ compiles all the workflow execution logs in one central location, 
making it easier to check any standard output (`stdout`) or standard error 
(`stderr`) for any command-line tools used on the workflow or the R code `stdout`.
Also, the workflow plot is appended at the beginning of the report, making it 
easier to click on the respective step.

```{r, eval=FALSE}
sal <- renderLogs(sal)
```

# Exported the workflow 

_`systemPipeR`_ workflow management system allows to translate and export the 
workflow build interactively to R Markdown format or an executable bash script.
This feature advances the reusability of the workflow, as well as the flexibility
for workflow execution.

## R Markdown file

`sal2rmd` function takes an `SYSargsList` workflow container and translates it to 
SPR workflow template R markdown format. This file can be imported with the 
`importWF` function, as demonstrated above.

```{r, eval=FALSE}
sal2rmd(sal)
```

## Bash script

`sal2bash` function takes an `SYSargsList` workflow container and translates 
it to an executable bash script, so one can run the workflow without loading 
`SPR` or using an R console.


```{r, eval=FALSE}
sal2bash(sal)
```

It will be generated on the project root an executable bash script, called by
default the `spr_wf.sh`. Also, a directory `./spr_wf` will be created and store 
all the R scripts based on the workflow steps. Please note that this function will 
"collapse" adjacent R steps into one file as much as possible.

# Project Resume and Restart

If you desire to resume or restart a project that has been initialized in the past, 
`SPRproject` function allows this operation.

With the `resume` option, it is possible to load the `SYSargsList` object in R and 
resume the analysis. Please, make sure to provide the `logs.dir` location, and the 
corresponded `YAML` file name, if the default names were not used when the project was created.

```{r SPR_resume, eval=FALSE}
sal <- SPRproject(resume = TRUE, logs.dir = ".SPRproject", 
                  sys.file = ".SPRproject/SYSargsList.yml") 
```

If you choose to save the environment in the last analysis, you can recover all 
the files created in that particular section. `SPRproject` function allows this 
with `load.envir` argument. Please note that the environment was saved only with
you run the workflow in the last section (`runWF()`).

```{r resume_load, eval=FALSE}
sal <- SPRproject(resume = TRUE, load.envir = TRUE) 
```

After loading the workflow at your current section, you can check the objects 
created in the old environment and decide if it is necessary to copy them to the
current environment.

```{r envir, eval=FALSE}
viewEnvir(sal)
copyEnvir(sal, list="plot", new.env = globalenv())
```

The `resume` option will keep all previous logs in the folder; however, if you desire to 
clean the execution (delete all the log files) history and restart the workflow, 
the `restart=TRUE` option can be used.

```{r restart_load, eval=FALSE}
sal <- SPRproject(restart = TRUE, load.envir = FALSE) 
```

The last and more drastic option from `SYSproject` function is to `overwrite` the
logs and the `SYSargsList` object. This option will delete the hidden folder and the 
information on the `SYSargsList.yml` file. This will not delete any parameter
file nor any results it was created in previous runs. Please use with caution. 

```{r SPR_overwrite, eval=FALSE}
sal <- SPRproject(overwrite = TRUE) 
```

# Exploring workflow instances {#sysargslist}

_`systemPipeR`_ provide several accessor methods and useful functions to explore 
`SYSargsList` workflow object.

## Accessor Methods

Several accessor methods are available that are named after the slot names of
the `SYSargsList` workflow object.

```{r}
names(sal)
```

- Check the length of the workflow:

```{r}
length(sal)
```

- Check the steps of the workflow:

```{r}
stepsWF(sal)
```

- Checking the command-line for each target sample:

`cmdlist()` method printing the system commands for running command-line
software as specified by a given `*.cwl` file combined with the paths to the
input samples (*e.g.* FASTQ files) provided by a `targets` file. The example below
shows the `cmdlist()` output for running `gzip` and `gunzip` on the first sample. 
Evaluating the output of `cmdlist()` can be very helpful for designing
and debugging `*.cwl` files of new command-line software or changing the
parameter settings of existing ones.

```{r}
cmdlist(sal, step = c(2,3), targets = 1)
```

- Check the workflow status:

```{r}
statusWF(sal)
```

- Check the workflow targets files:

```{r}
targetsWF(sal[2])
```

- Checking the expected outfiles files:

The `outfiles` components of `SYSargsList` define the expected outfiles files 
for each step in the workflow, some of which are the input for the next workflow step.

```{r}
outfiles(sal[2])
```

- Check the workflow dependencies:

```{r}
dependency(sal)
```

- Check the sample comparisons:

Sample comparisons are defined in the header lines of the `targets` file 
starting with '``# <CMP>``'. This information can be accessed as follows:

```{r, eval=FALSE}
targetsheader(sal, step = "Quality")
```

- Get the workflow steps names: 

```{r}
stepName(sal)
```

- Get the Sample Id for on particular step:

```{r}
SampleName(sal, step = "gzip")
SampleName(sal, step = "iris_stats")
```

- Get the `outfiles` or `targets` column files: 

```{r}
getColumn(sal, "outfiles", step = "gzip", column = "gzip_file")
getColumn(sal, "targetsWF", step = "gzip", column = "FileName")
```

- Get the R code for a `LineWise` step:

```{r}
codeLine(sal, step = "export_iris")
```

- View all the objects in the running environment:

```{r}
viewEnvir(sal)
```

- Copy one or multiple objects from the running environment to a new environment:

```{r}
copyEnvir(sal, list = c("plot"), new.env = globalenv(), silent = FALSE)
```

- Accessing the `*.yml` data

```{r}
yamlinput(sal, step = "gzip")
```

## Subsetting the workflow details

- The `SYSargsList` class and its subsetting operator `[`:

```{r}
sal[1]
sal[1:3]
sal[c(1,3)]
```

- The `SYSargsList` class and its subsetting by steps and input samples: 

```{r, eval=FALSE}
sal_sub <- subset(sal, subset_steps = c( 2,3), input_targets = ("SE"), keep_steps = TRUE)
stepsWF(sal_sub)
targetsWF(sal_sub)
outfiles(sal_sub)
```

- The `SYSargsList` class and its operator `+`

```{r, eval=FALSE}
sal[1] + sal[2] + sal[3]
```

## Replacement Methods

- Update a `input` parameter in the workflow

```{r, eval=FALSE}
sal_c <- sal
## check values
yamlinput(sal_c, step = "gzip")
## check on command-line
cmdlist(sal_c, step = "gzip", targets = 1)
## Replace
yamlinput(sal_c, step = "gzip", paramName = "ext") <- "txt.gz"

## check NEW values
yamlinput(sal_c, step = "gzip")
## Check on command-line
cmdlist(sal_c, step = "gzip", targets = 1)
```

- Append and Replacement methods for R Code Steps

```{r, sal_lw_rep, eval=FALSE}
appendCodeLine(sal_c, step = "export_iris", after = 1) <- "log_cal_100 <- log(100)"
codeLine(sal_c, step = "export_iris")

replaceCodeLine(sal_c, step="export_iris", line = 2) <- LineWise(code={
                    log_cal_100 <- log(50)
                    })
codeLine(sal_c, step = 1)
```

For more details about the `LineWise` class, please see [below](#linewise).

- Rename a Step

```{r, eval=FALSE}
renameStep(sal_c, step = 1) <- "newStep"
renameStep(sal_c, c(1, 2)) <- c("newStep2", "newIndex")
sal_c
names(outfiles(sal_c))
names(targetsWF(sal_c))
dependency(sal_c)
```

- Replace a Step

```{r, eval=FALSE}
sal_test <- sal[c(1,2)]
replaceStep(sal_test, step = 1, step_name = "gunzip" ) <- sal[3]
sal_test
```

Note: Please use this method with attention, because it can disrupt all 
the dependency graphs. 

- Removing a Step

```{r}
sal_test <- sal[-2]
sal_test
```


# Session 
```{r}
sessionInfo()
```

# Sub-sections
