---
title: "Design and run Workflows" 
author: "Author: Daniela Cassol (danielac@ucr.edu) and Thomas Girke (thomas.girke@ucr.edu)"
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`" 
output:
  BiocStyle::html_document:
    toc_float: true
    code_folding: show
  BiocStyle::pdf_document: default
package: systemPipeR
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{systemPipeR: Workflow design and reporting generation environment}
  %\VignetteEngine{knitr::rmarkdown}
fontsize: 14pt
bibliography: bibtex.bib
editor_options: 
  chunk_output_type: console
type: docs
weight: 3
---

<!--
- Compile from command-line
Rscript -e "rmarkdown::render('systemPipeR.Rmd', c('BiocStyle::html_document'), clean=F); knitr::knit('systemPipeR.Rmd', tangle=TRUE)"; Rscript ../md2jekyll.R systemPipeR.knit.md 2; Rscript -e "rmarkdown::render('systemPipeR.Rmd', c('BiocStyle::pdf_document'))"
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
options(width=80, max.print=1000)
knitr::opts_chunk$set(
    eval=as.logical(Sys.getenv("KNITR_EVAL", "TRUE")),
    cache=as.logical(Sys.getenv("KNITR_CACHE", "TRUE")), 
    tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
    library(systemPipeR)
})
```

```{r eval=TRUE, include=FALSE}
# cleaning
try(unlink(".SPRproject", recursive = TRUE), TRUE)
try(unlink("data", recursive = TRUE), TRUE)
try(unlink("results", recursive = TRUE), TRUE)
try(unlink("param", recursive = TRUE), TRUE)
```

# About this section

In this section, we will discuss following topics: 

- How to create SPR data analysis projects.
- How to build workflow step by step interactively or use use a template as starting point.
- After step design, how to run a workflow.
- After workflow finished running, how we can check the status, visualize the workflow.
- Different options for managing the workflow, _e.g._ resume, restart, overwrite a 
  SPR project. 
- How to explore the workflow object (methods).
- Finally, how to generate some data analysis reports.


# Project initialization

To create a workflow within _`systemPipeR`_, we can start by defining an empty
container and checking the directory structure:

```{r SPRproject1, eval=TRUE}
sal <- SPRproject() 
```

Internally, `SPRproject` function will create a hidden folder called `.SPRproject`,
by default, to store all the log files.
A `YAML` file, here called `SYSargsList.yml`, has been created, which initially
contains the basic location of the project structure; however, every time the 
workflow object `sal` is updated in R, the new information will also be store in this 
flat-file database for easy recovery.
If you desire different names for the logs folder and the `YAML` file, these can 
be modified as follows:

```{r SPRproject_logs, eval=FALSE}
sal <- SPRproject(logs.dir= ".SPRproject", sys.file=".SPRproject/SYSargsList.yml") 
```

Also, this function will check and/or create the basic folder structure if missing, 
which means `data`, `param`, and `results` folder, as described [here](https://systempipe.org/sp/spr/gettingstarted/#directory-structure). 
If the user wants to use a different names for these directories, can be specified 
as follows:

```{r SPRproject_dir, eval=FALSE}
sal <- SPRproject(data = "data", param = "param", results = "results") 
```

It is possible to separate all the R objects created within the workflow analysis 
from the current environment. `SPRproject` function provides the option to create 
a new environment, and in this way, it is not overwriting any object you may want
to have at your current section. 

```{r SPRproject_env, eval=FALSE}
sal <- SPRproject(envir = new.env()) 
```

In this stage, the object `sal` is a empty container, except for the project information. The project information can be accessed by the `projectInfo` method:

```{r projectInfo, eval=TRUE}
sal
projectInfo(sal)
```

Also, the `length` function will return how many steps this workflow contains, 
and in this case, it is empty, as follow:

```{r length, eval=TRUE}
length(sal)
```

# Workflow Design 

_`systemPipeR`_ workflows can be designed and built from start to finish with a single command, importing from an R Markdown file or stepwise in interactive mode from the R console. 
In the [next section](#appendstep), we will demonstrate how to build the workflow in an interactive mode, and in the [following section](#importWF), we will show how to build from a file. 

New workflows are constructed, or existing ones modified, by connecting each step
via `appendStep` method. Each `SYSargsList` instance contains instructions needed 
for processing a set of input files with a specific command-line and the paths to 
the corresponding outfiles generated.

The constructor function `Linewise` is used to build the R code-based step. 
For more details about this S4 class container, see [here](#linewise). 

## Build workflow interactive {#appendstep}

This tutorial shows a straightforward example for describing and explaining all main features available within systemPipeR to design, build, manage, run, and visualize the workflow. In summary, we are exporting a dataset to multiple files, compressing and decompressing each one of the files, importing to R, and finally performing a statistical analysis. 

In the previous section, we initialize the project by building the `sal` object.
Until this moment, the container has no steps:

```{r sal_check, eval=TRUE}
sal
```

In the next subsection, we will discuss how to populate the object created with the first step in the
workflow interactively.

## Subsections

```{r eval=TRUE, include=FALSE}
# cleaning
try(unlink(".SPRproject", recursive = TRUE), TRUE)
try(unlink("data", recursive = TRUE), TRUE)
try(unlink("results", recursive = TRUE), TRUE)
try(unlink("param", recursive = TRUE), TRUE)
```
