---
title: "Export workflows" 
date: "Last update: `r format(Sys.time(), '%d %B, %Y')`" 
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{systemPipeR: Workflow design and reporting generation environment}
  %\VignetteEngine{knitr::rmarkdown}
fontsize: 14pt
editor_options: 
  chunk_output_type: console
type: docs
weight: 8
---

_`systemPipeR`_ workflow management system allows one to translate workflow into 
**reproducible code** and export the 
workflow build interactively to R Markdown template format or an executable bash script.
This feature advances the reusability of the workflow, as well as the flexibility
for workflow execution.



```{r eval=TRUE, include=FALSE}
# cleaning
try(unlink(".SPRproject", recursive = TRUE), TRUE)
try(unlink("data", recursive = TRUE), TRUE)
try(unlink("results", recursive = TRUE), TRUE)
try(unlink("param", recursive = TRUE), TRUE)
try(unlink("spr_template.md", recursive = TRUE), TRUE)
try(unlink("spr_template.Rmd", recursive = TRUE), TRUE)
try(unlink("spr_bash", recursive = TRUE), TRUE)
try(unlink("spr_wf.sh", recursive = TRUE), TRUE)
```



```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
    library(systemPipeR)
})
```

We still use the [simple workflow](https://raw.githubusercontent.com/systemPipeR/systemPipeR.github.io/main/static/en/sp/spr/sp_run/spr_simple_wf.md) to demonstrate.
```{r collapse=TRUE}
sal <- SPRproject()
sal <- importWF(sal, file_path = system.file("extdata", "spr_simple_wf.Rmd", package = "systemPipeR"), verbose = FALSE)
sal
```

## R Markdown file

`sal2rmd` function takes an `SYSargsList` workflow container and translates it to 
SPR workflow template R markdown format. This file can be imported with the 
[`importWF` function](../step_import). This is very similar to the 
[text-enriched report](../step_reports), but with all text removed. 

```{r, collapse=TRUE}
sal2rmd(sal)
```


## Bash script

`sal2bash` function takes an `SYSargsList` workflow container and translates 
it to an executable bash script, The benefit of this is that if the workflow is 
only composed with command-line steps, no R step (LineWise) step in involved, then 
one could generate the workflow on one computer and run in another computer without 
installing SPR or R at all. 

```{r, collapse=TRUE}
sal2bash(sal)
```

It will be generated on the project root an executable bash script, called by
default the `spr_wf.sh`. Also, a directory `./spr_wf` will be created and store 
all the R scripts based on the workflow steps. Please note that this function will 
"collapse" adjacent R steps into one file as much as possible.

```{r}
fs::dir_tree("spr_bash")
```


# Session 
```{r}
sessionInfo()
```

```{r eval=TRUE, include=FALSE}
# cleaning
try(unlink(".SPRproject", recursive = TRUE), TRUE)
try(unlink("data", recursive = TRUE), TRUE)
try(unlink("results", recursive = TRUE), TRUE)
try(unlink("param", recursive = TRUE), TRUE)
try(unlink("spr_template.Rmd", recursive = TRUE), TRUE)
try(unlink("spr_template.md", recursive = TRUE), TRUE)
try(unlink("spr_bash", recursive = TRUE), TRUE)
try(unlink("spr_wf.sh", recursive = TRUE), TRUE)
```
