---
title: "Import workflow from template" 
fontsize: 14pt
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 80
type: docs
weight: 2
---

## Build workflow from a template

The precisely [same workflow](../step_interactive) can be created by importing
the steps from a template file. In SPR, we use R Markdown files as templates. As
demonstrated above, it is required to initialize the project with `SPRproject`
function.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
    library(systemPipeR)
})
```

### `importWF`

`importWF` function will scan and import all the R chunk from the R Markdown
file and build all the workflow instances. Then, each R chuck in the file will
be converted in a workflow step.

We have prepared the template for you already. Let's first see what is in the
template:

```{r comment=""}
cat(readLines(system.file("extdata", "spr_simple_wf.Rmd", package = "systemPipeR")), sep = "\n")
```


#### SPR chunks 
The SPR templates has no difference than a normal R markdown file, except one 
little thing -- the **SPR chunks**.

To make a normal R chunk also a SPR chunk, in the chunk header `spr=TRUE` or 
`spr=T` option needs to be appended. 

For example:

> *\`\`\`{r step_1, eval=TRUE, spr=TRUE}*

> *\`\`\`{r step_2, eval=FALSE, spr=TRUE}*

Note here the `eval=FALSE`, by default steps with this option will still 
be imported, but you can use `ignore_eval` flag to change it in `importWF`. 

##### Preprocess code 
Inside SPR chunks, before the actual step definition, there is some special space called 
preprocess code. 

Why do need preprocess code? When we import/create the workflow steps, these 
steps are not really executed when the time of creation, no matter it is a `sysArgs`
step or a `Linewise` step. However, in many cases, we need to connect different previous steps'
outputs to the inputs of the next.  This is easy to handle between `sysArgs` steps via the `targets` 
argument connection. If it is a `Linewise` step to a `sysArgs` step. Things become
tricky. Since `Linewise` code is not run at the time of step definition, no 
output paths are generated, so the next  `sysArgs` step cannot find the inputs.

To overcome this problem, preprocess code feature is introduced. Defining preprocess
code is very easy. Write any lines of R code below the SPR chunk header line. Right 
before the step is defined, insert one line of comment of `###pre-end` to indicate
the completion of preprocess code. For example:

```{r eval=FALSE}
targetspath <- system.file("extdata", "cwl", "example", "targets_example.txt", package = "systemPipeR")

###pre-end
appendStep(sal) <- SYSargsList(
  step_name = "Example", 
  targets = targetspath, 
  wf_file = "example/example.cwl", input_file = "example/example.yml", 
  dir_path = system.file("extdata/cwl", package = "systemPipeR"), 
  inputvars = c(Message = "_STRING_", SampleName = "_SAMPLE_")
)
```

In the example above the targets path is not directly loaded but given through an intermediate 
variable `targetspath`. This is a simple example, other useful actions like path 
concatenation, checking file integrity before piping to expensive (slow) functions
can also be done in preprocess.


Actually, if the SPR chunk has R code before the step definition but `###pre-end`
delimiter is not added, these code will **still** be evaluated at the time of import. 
However, these lines of code will not be store in the `SYSargsList`, so later 
when you render the report (`renderReport`) or export the workflow as a new template 
(`sal2rmd`), these lines will **not** be included. That means, these lines are 
**one-shot** reprocess code and not reproducible. 

##### Other rules

- For SPR chunks, the last object assigned must to be the `SYSargsList`, for
example a `sysArgs2`(commandline) steps:

    ```{r fromFile_example_rules_cmd, eval=FALSE}
    targetspath <- system.file("extdata/cwl/example/targets_example.txt", package = "systemPipeR")
    appendStep(sal) <- SYSargsList(step_name = "Example", 
                          targets = targetspath, 
                          wf_file = "example/example.cwl", input_file = "example/example.yml", 
                          dir_path = system.file("extdata/cwl", package = "systemPipeR"), 
                          inputvars = c(Message = "_STRING_", SampleName = "_SAMPLE_"))
    ```

    OR a `Linewise` (R) step:
    
    ```{r fromFile_example_rules_r, eval=FALSE}
    appendStep(sal) <- LineWise(code = {
                                  library(systemPipeR)
                                },
                                step_name = "load_lib")
    ```

- Also, note that all the required files or objects to generate one particular
step must be defined in an imported R code. The motivation for this is that when
R Markdown files are imported, the `spr = TRUE` flag will be evaluated, append,
and stored in the workflow control class as the `SYSargsList` object.

- The workflow object name used in the R Markdown (e.g. `appendStep(sal)`) needs
to be the same when the `importWF` function is used. Usually we use the name `sal` ( 
short abbreviation for `sysargslist`). It is important to keep
consistency. If different object names are used, when running the workflow, you
can see a error, like *Error:* <objectName> object not found..

- Special in `importWF`: SPR chunk names will be used as step names, and it has 
  higher priority than the `stepname` argument. For example, the chunk header is 
  `{r step_1, eval=TRUE, spr=TRUE}`, and the inside `SYSargsList` option is 
  `SYSargsList(step_name = "step_99", ...)`. After the import, step name will be 
  overwritten to `"step_1"` instead of `"step_99"`.
  

#### start to import

```{r cleaning3, eval=TRUE, include=FALSE}
if (file.exists(".SPRproject_rmd")) unlink(".SPRproject_rmd", recursive = TRUE)
```

```{r importWF_rmd, eval=TRUE, collapse=TRUE}
sal_rmd <- SPRproject(logs.dir = ".SPRproject_rmd") 

sal_rmd <- importWF(sal_rmd, file_path = system.file("extdata", "spr_simple_wf.Rmd", package = "systemPipeR"))
```


Let's explore the workflow to check the steps:

```{r importWF_details}
stepsWF(sal_rmd)
dependency(sal_rmd)
codeLine(sal_rmd)
targetsWF(sal_rmd)
```


# Session

```{r}
sessionInfo()
```
